---
title: "Figure 5 and associated supplementary figures"
output: html_notebook
author: Grace Avecilla
---

What is the best way to use aggregate observations? Single inference to distribution (WF + NPE) 

```{r libs}
library(tidyverse)
library(patchwork)
library(scales)
library(ggbeeswarm)
library(reticulate)
theme_set(theme_classic(base_size = 21))
file_dir = "/Volumes/GoogleDrive/.shortcut-targets-by-id/19NQRq-XLPpjqJGmLbP4MHRjala2WVahk/Avecilla et al Simulating CNV evolution/scripts/cnv_sims_inference/hpc_output/test_performance/multi/"
fig_path = "/Volumes/GoogleDrive/.shortcut-targets-by-id/19NQRq-XLPpjqJGmLbP4MHRjala2WVahk/Avecilla et al Simulating CNV evolution/figures/"

```

```{r functions}
read_est_params = function(x) {
  if(str_detect(x, "NPE")==TRUE){
    file = read_csv(x, col_names = c("Model","Method","col_to_sep",
                                     "s_est","mu_real",
                                     "mu_est","s_snv","m_snv","rmse_map",
                                     "mean_rmse_ppc","rmse95_low_ppc","rmse95_hi_ppc",
                                     "corr_map",
                                     "mean_corr_ppc","corr95_low_ppc", "corr95_hi_ppc",
                                     "fit_95hdi_low","fit_95hdi_high",
                                     "mut_95hdi_low","mut_95hdi_high",
                                     "aic", "dic", "waic1", "waic2")) %>%
      mutate(n_sim=NA) %>%
      separate(col_to_sep, into = c(NA,NA,NA,"n_presim", "presim_rep","s_real"), 
               sep="([_.])") %>%
      mutate(s_real = str_remove(s_real, "csv")) %>%
      mutate(s_est_10 = s_est, s_real_10 = as.numeric(s_real),
             mu_est_10 = mu_est, mu_real_10 = mu_real,
             s_est = 10^s_est, s_real = 10^as.numeric(s_real),
             mu_est = 10^mu_est, mu_real = 10^mu_real) %>%
      mutate(mu_ratio = mu_est/mu_real, s_ratio = s_est/s_real,
             mu_diff = mu_est - mu_real, s_diff = s_est - s_real) %>%
      mutate(flow_type = if_else(str_detect(x, "nsf"), "NSF", "MAF")) 
  }
  if(str_detect(x, "ABC")==TRUE) {
    file = read_csv(x, col_names = c("Model","Method","col_to_sep",
                                     "s_est","mu_real",
                                     "mu_est","s_snv","m_snv","rmse_map",
                                     "mean_rmse_ppc","rmse95_low_ppc","rmse95_hi_ppc",
                                     "corr_map",
                                     "mean_corr_ppc","corr95_low_ppc", "corr95_hi_ppc",
                                     "fit_95hdi_low","fit_95hdi_high",
                                     "mut_95hdi_low","mut_95hdi_high",
                                     "aic", "dic", "waic1", "waic2", "n_sim")) %>%
      separate(col_to_sep, into = c("n_presim","s_real"), 
               sep="([-])") %>%
      mutate(s_real = as.numeric(s_real)*-1) %>%
      mutate(presim_rep = case_when(str_detect(x, "_1_") ~ "1",
                                    str_detect(x, "_2_") ~ "2",
                                    str_detect(x, "_3_") ~ "3")) %>%
      mutate(n_sim = if("n_sim" %in% colnames(.)) n_sim else NA) %>%
      select("Model","Method","n_presim", "presim_rep","s_real",
             "s_est","mu_real",
             "mu_est","s_snv","m_snv","rmse_map",
             "mean_rmse_ppc","rmse95_low_ppc","rmse95_hi_ppc",
             "corr_map",
             "mean_corr_ppc","corr95_low_ppc", "corr95_hi_ppc",
             "fit_95hdi_low","fit_95hdi_high",
             "mut_95hdi_low","mut_95hdi_high",
             "aic", "dic", "waic1", "waic2", "n_sim") %>%
      mutate(s_est_10 = s_est, s_real_10 = as.numeric(s_real),
             mu_est_10 = mu_est, mu_real_10 = mu_real,
             s_est = 10^s_est, s_real = 10^as.numeric(s_real),
             mu_est = 10^mu_est, mu_real = 10^mu_real) %>%
      mutate(mu_ratio = mu_est/mu_real, s_ratio = s_est/s_real,
             mu_diff = mu_est - mu_real, s_diff = s_est - s_real) %>%
      mutate(flow_type = NA) %>%
      mutate(n_presim = if_else(n_presim == "100", "10000", "100000"))
  }
  return(file)
}

read_est_params_single = function(x) {
  data = read_est_params(x)
  data = data %>% 
    mutate(obs_set = str_extract(x, pattern="\\w+.([0-1]+)_([mut57]+)")) %>%
    select(-n_sim)
  return(data)
}

read_est_params_multi = function(x) {
  if(str_detect(x, "NPE")==TRUE){
    file = read_csv(x, col_names = c("obs_set", "held_out","Model","Method","col_to_sep",
                                     "shape_est","scale_real","scale_est","mu_real",
                                     "mu_est","s_snv","m_snv",
                                     "mean_rmse_ppc","rmse95_low_ppc","rmse95_hi_ppc",
                                     "mean_corr_ppc","corr95_low_ppc", "corr95_hi_ppc",
                                     "shape_95hdi_low","shape_95hdi_high",
                                     "scale_95hdi_low","scale_95hdi_high",
                                     "mut_95hdi_low","mut_95hdi_high",
                                     "aic", "dic", "waic1", "waic2",
                                     "var_observation", "var_ppc")) %>%
      separate(col_to_sep, into = c(NA,NA,NA,NA,"n_presim", "presim_rep","shape_real"), 
               sep="([_.])") %>%
      mutate(shape_real = as.numeric(str_remove(shape_real, "csv"))) %>%
      mutate(scale_est_10 = scale_est, scale_real_10 = scale_real,
             mu_est_10 = mu_est, mu_real_10 = mu_real,
             scale_est = 10^scale_est, scale_real = 10^scale_real,
             mu_est = 10^mu_est, mu_real = 10^mu_real) %>%
      mutate(flow_type = if_else(str_detect(x, "nsf"), "NSF", "MAF"))
  }
  if(str_detect(x, "ABC")==TRUE) {
    file = read_csv(x, col_names = c("obs_set", "held_out","Model","Method","col_to_sep",
                                     "shape_est","scale_real","scale_est","mu_real",
                                     "mu_est","s_snv","m_snv",
                                     "mean_rmse_ppc","rmse95_low_ppc","rmse95_hi_ppc",
                                     "mean_corr_ppc","corr95_low_ppc", "corr95_hi_ppc",
                                     "shape_95hdi_low","shape_95hdi_high",
                                     "scale_95hdi_low","scale_95hdi_high",
                                     "mut_95hdi_low","mut_95hdi_high",
                                     "aic", "dic", "waic1", "waic2",
                                     "var_observation", "var_ppc")) %>%
      mutate(n_presim = if_else(col_to_sep%%2 == 0, 
                                as.numeric(str_sub(as.character(col_to_sep), end=-3)),
                                as.numeric(str_sub(as.character(col_to_sep), end=-2)))) %>%
      mutate(shape_real = if_else(col_to_sep%%2 == 0, 10,1)) %>%
      mutate(presim_rep = case_when(str_detect(x, "_1_") ~ "1",
                                    str_detect(x, "_2_") ~ "2",
                                    str_detect(x, "_3_") ~ "3")) %>%
      select("obs_set", "held_out","Model","Method","n_presim", "presim_rep","shape_real",
             "shape_est","scale_real","scale_est","mu_real",
             "mu_est","s_snv","m_snv",
             "mean_rmse_ppc","rmse95_low_ppc","rmse95_hi_ppc",
             "mean_corr_ppc","corr95_low_ppc", "corr95_hi_ppc",
             "shape_95hdi_low","shape_95hdi_high",
             "scale_95hdi_low","scale_95hdi_high",
             "mut_95hdi_low","mut_95hdi_high",
             "aic", "dic", "waic1", "waic2","var_observation", "var_ppc") %>%
      mutate(scale_est_10 = scale_est, scale_real_10 = scale_real,
             mu_est_10 = mu_est, mu_real_10 = mu_real,
             scale_est = 10^scale_est, scale_real = 10^scale_real,
             mu_est = 10^mu_est, mu_real = 10^mu_real) %>%
      mutate(flow_type = "") %>%
      mutate(n_presim = if_else(n_presim == "100", "10000", "100000"))
  }
  return(file %>% mutate(obs_set = str_extract(obs_set, "\\w+.([0-1]+)_([mut57]+)")))
}

```

```{r get data}
files = list.files(path = paste0(file_dir,"param_estimates_single_multi/"), pattern = '*est_real_params.csv')
files = paste0(file_dir, "param_estimates_single_multi/", files)
data_list = map(files,read_est_params_single)
single_data_all = do.call(rbind, data_list)

files = list.files(path = paste0(file_dir,"param_estimates_multi_multi/"), pattern = '*est_real_params.csv')
files = paste0(file_dir, "param_estimates_multi_multi/", files)
data_list = map(files,read_est_params_multi)
multi_data_all = do.call(rbind, data_list)
```

```{r}
single_data = single_data_all %>% filter(n_presim == "100000" & flow_type == "NSF" & Model == "WF") %>% 
  unite(mod_method, Model, Method, sep = " ", remove=F) %>%
  unite(sim_params,s_real, mu_real, remove = F) %>%
  mutate(alpha_real = if_else(str_detect(obs_set,"shape1_"), 1, 10),
         beta_real = case_when(str_detect(obs_set, "scale0.001")~ 0.001,
                               str_detect(obs_set, "scale0.0001")~ 0.0001,
                               str_detect(obs_set, "scale0.01")~ 0.01,
                               str_detect(obs_set, "scale0.1")~ 0.1))

multi_data = multi_data_all %>%filter(n_presim == "100000" & flow_type == "NSF" & Model == "WF") %>%
  mutate(es = shape_real*scale_real) %>%
  unite(sim_params,es, mu_real, remove = F) %>%
  mutate(es_est = shape_est*scale_est)

color_vec = c("#4daf4a", "#984ea3","#e41a1c", "#377eb8", "black")
mod_meth = c("Chemo NPE", "WF NPE", "Chemo ABC-SMC", "WF ABC-SMC", "Pseudo-observation")

obs_names = unique(single_data$obs_set)

```

```{python}
import scipy.stats
import scipy.special
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt

leave_outs = np.array([[0,3,6],[1,4,7],[2,5,8]])
out = []
for i in r.obs_names:
        data=r.single_data.loc[r.single_data['obs_set'] == i]
        data = pd.DataFrame.to_numpy(data['s_est'])
        for j in leave_outs:
                new_data = np.delete(data, j)
                fit_alpha, fit_loc, fit_beta = scipy.stats.gamma.fit(new_data, floc=0)
                t=[fit_alpha, fit_loc, fit_beta, i]
                out.append(t)
single_inferred_dfe = pd.DataFrame(out, columns = ['single_alpha', 'single_loc', 'single_beta', 'obs_set'])

```

```{python}
#calculate kl divergence
out = []
for i in r.obs_names:
        data_s=single_inferred_dfe.loc[single_inferred_dfe['obs_set'] == i]
        data_m = r.multi_data.loc[r.multi_data['obs_set'] == i]
        alpha_real = data_m['shape_real'].iloc[0]
        beta_real = data_m['scale_real'].iloc[0]
        mut_real = data_m['mu_real'].iloc[0]
        for j in range(3):
                 x=np.linspace(beta_real*alpha_real/10,beta_real*alpha_real*2, 100)
                 p=scipy.stats.gamma.pdf(x, alpha_real, 0, beta_real)
                 q=scipy.stats.gamma.pdf(x, data_s['single_alpha'].iloc[j], 0, data_s['single_beta'].iloc[j])
                 p[np.where(p == 0)[0]] = 1e-10
                 q[np.where(q == 0)[0]] = 1e-10
                 kls=np.ma.masked_invalid(scipy.special.kl_div(p, q)).sum()
                 q=scipy.stats.gamma.pdf(x, data_m['shape_est'].iloc[j], 0, data_m['scale_est'].iloc[j])
                 klm=np.ma.masked_invalid(scipy.special.kl_div(p, q)).sum()
                 t=[kls, klm, i, alpha_real, beta_real, mut_real]
                 out.append(t)
kl_div = pd.DataFrame(out, columns = ['kl_div_single', 'kl_div_multi', 'obs_set', 'alpha_real', 'beta_real', 'mu_real'])
kl_div.round({'kl_div_single': 1, 'kl_div_multi': 2}).to_csv("SupplementaryTable1.csv")
```

```{r}
multi_data$sim_params = factor(multi_data$sim_params, levels = c("0.001_1e-05", "0.001_1e-07", "0.1_1e-05", "0.1_1e-07"),
                         ordered = TRUE, labels=c(expression(atop("s = 0.001,", paste(delta, "  = 1e-5"))),
                                                  expression(atop("s = 0.001,", paste(delta, "  = 1e-7"))),
                                                  "0.1_1e-05"  = expression(atop("s = 0.1,", paste(delta, "  = 1e-5"))),
                                                  "0.1_1e-07"  = expression(atop("s = 0.1,", paste(delta, "  = 1e-7")))))
```

```{r curves f5}
plot_dist = function(x, legend=TRUE) {
  temp = multi_data %>%
    filter(obs_set == x)
  temp2 = py$single_inferred_dfe %>%
    filter(obs_set == x)
  if(temp$es[1] == 0.1) {
    d = data.frame(x = seq(0,1, length.out = 100))
    bw=0.1
    y_ann = 7.5
    x_ann = 0.75
  }
  if(temp$es[1] == 0.001) {
    d = data.frame(x = seq(0,0.04, length.out =  100))
    bw=0.001
    y_ann=1000
    x_ann=0.03
  }
  p = ggplot(data = d, aes(x)) +
    #geom_histogram(data = temp2, aes(x = value, fill = s_type), alpha = 0.5, binwidth=bw) +
    stat_function(fun = function(x) 
           {dgamma(x, shape = temp$shape_real[1], scale = temp$scale_real[1])},# * 20 * bw}, 
           aes(colour = "True DFE"), size = 2, alpha=1) +
    stat_function(fun = function(x) 
           {dgamma(x, shape = temp$shape_est[1], scale = temp$scale_est[1])},# * 20 * bw}, 
           aes(colour = "Observation set DFE"), size = 1, alpha = 0.6) +
    stat_function(fun = function(x) 
           {dgamma(x, shape = temp$shape_est[2], scale = temp$scale_est[2])},# * 20 * bw}, 
           aes(colour = "Observation set DFE"), size = 1, alpha = 0.6) +
    stat_function(fun = function(x) 
           {dgamma(x, shape = temp$shape_est[3], scale = temp$scale_est[3])},# * 20 * bw}, 
           aes(colour = "Observation set DFE"), size = 1, alpha = 0.6) +
    #####
      stat_function(fun = function(x) 
           {dgamma(x, shape = temp2$single_alpha[1], scale = temp2$single_beta[1])},# * 20 * bw}, 
           aes(colour = "Single observation DFE"), size = 1, alpha = 0.6) +
    stat_function(fun = function(x) 
           {dgamma(x, shape = temp2$single_alpha[2], scale = temp2$single_beta[2])},# * 20 * bw}, 
           aes(colour = "Single observation DFE"), size = 1, alpha = 0.6) +
    stat_function(fun = function(x) 
           {dgamma(x, shape = temp2$single_alpha[3], scale = temp2$single_beta[3])},# * 20 * bw}, 
           aes(colour = "Single observation DFE"), size = 1, alpha = 0.6) +
    scale_color_manual(breaks = c("True DFE","Observation set DFE",
                                  "Single observation DFE"), name = "",
                     #values=c("#453781", "#1f968b", "#75d054")) +
                     values=c("black", "darkgreen", "blue")) +
    xlab(expression(paste("selection coefficient, ",italic("s"[C])))) +
    ylab("density") +
    annotate("text",label=as.expression(bquote(atop("E(s)="~.(temp$es[1])~",",delta[C]~"= "~.(temp$mu_real[1])))),
            x=x_ann, y = y_ann, size=6)
  if(legend == TRUE) {
    return(p)
  } 
  else {
    return(p + theme(legend.position = "none"))
  }
}

plot_list = list()
i=1
al = c("A", "B","C","D","A", "B","C","D")
for(obs in unique(single_data$obs_set)) {
  #print(plot_dist(obs))
  plot_list[[i]] = plot_dist(obs, legend = FALSE) + ggtitle(al[i])
  i=i+1
}
plot_list[[4]] = plot_dist(unique(single_data$obs_set)[4], legend = TRUE) + ggtitle(al[4])
plot_list[[8]] = plot_dist(unique(single_data$obs_set)[8], legend = TRUE) + ggtitle(al[8])

```

```{r insets f5}
plot_kldiv = function(x) {
  temp = as_tibble(py$kl_div) %>%
    filter(obs_set == x) 
  p=temp %>%
    pivot_longer(starts_with("kl"), names_to = "type", values_to = "KL divergence") %>%
    ggplot(aes(type, `KL divergence`)) +
    geom_beeswarm(size = 3) +
    scale_x_discrete(labels = c("Directly inferred\nMAP distribution","Distribution from\nsingle observation MAPs"))
  return(p)
}

inset_list = list()
i=1
for(obs in unique(single_data$obs_set)) {
  #print(plot_dist(obs))
  inset_list[[i]] = plot_kldiv(obs)
  i=i+1
}

plot_list[[1]] + inset_element(inset_list[[1]],0.4, 0.4, 1, 1, align_to = 'full')
```


```{r fig 5}
fig5 = plot_list[[1]] + plot_list[[2]] + plot_list[[3]] + plot_list[[4]]
fig5
ggsave(paste0(fig_path,"Figure5.pdf"), plot = fig5, width = 20, height = 10, units = "in")
ggsave(paste0(fig_path,"Figure5.png"), plot = fig5, width = 20, height = 10, units = "in")
```

*Figure 5.*  A set of eleven simulated synthetic observations was generated from a Wright-Fisher model with CNV selection coefficients sampled from an exponential distribution of fitness effects (DFE) (black curve). The MAP DFEs (blue curves) were directly inferred using three different subsets of eight out of eleven synthetic observations. We also inferred the selection coefficient for each observation in the set of eleven individually, and fit Gamma distributions to sets of eight inferred selection coefficients (green curves). All inferences were performed with NPE using the same amortized network to infer a posterior for each set of eight synthetic observations or each single observation. 


```{r sup fig 9}
supp = plot_list[[5]] + plot_list[[6]] + plot_list[[7]] + plot_list[[8]] 
supp
ggsave(paste0(fig_path,"SupplementaryFig9.pdf"), plot = supp, width = 20, height = 10, units = "in")
ggsave(paste0(fig_path,"SupplementaryFig9.png"), plot = supp, width = 20, height = 10, units = "in")
```

*Supplementary Figure 9.*  A set of eleven simulated synthetic observations was generated from a Wright-Fisher model with CNV selection coefficients sampled from an Gamma distribution where =10 of fitness effects (DFE) (black curve) . The MAP DFEs (blue curves) were directly inferred using three different subsets of eight out of eleven synthetic observations. We also inferred the selection coefficient for each observation in the set of eleven individually, and fit Gamma distributions to sets of eight inferred selection coefficients (green curves). All inferences were performed with NPE using the same amortized network to infer a posterior for each set of eight synthetic observations or each single observation.  

```{r sup fig 10}
a = multi_data %>%
  select(obs_set, shape_real, scale_real, es, mu_real, 
         mean_rmse_ppc, rmse95_low_ppc, rmse95_hi_ppc) %>%
  mutate(type = "multi") %>%
  bind_rows(single_data %>%
              rename(shape_real = alpha_real, scale_real = beta_real) %>%
              mutate(es = shape_real*scale_real, type = "single") %>%
              select(obs_set, shape_real, scale_real, es, mu_real,
                     mean_rmse_ppc, rmse95_low_ppc, rmse95_hi_ppc, type)) %>%
  mutate(sim_params = paste(es, mu_real, sep = "_")) %>%
  ggplot(aes(sim_params, mean_rmse_ppc, color = type)) +
  geom_pointrange(aes(ymin = rmse95_low_ppc,ymax = rmse95_hi_ppc), alpha=0.5,
                  position=position_dodge2(width=0.75)) +
  facet_wrap(~shape_real) +
  ylab("Posterior predictive check \nroot mean square error") +
  scale_color_manual(labels = c("Directly inferred MAP distribution",
                                  "Distribution from single observation MAPs"), name = "",
                     #values=c("#453781", "#1f968b", "#75d054")) +
                     values=c("blue", "darkgreen")) +
  ggtitle("A") +
  scale_x_discrete(name = "",
                   labels=c("0.001_1e-07"  = expression(atop("E(s) = 0.001,", paste(delta[C], "  = 1e-7"))),
                            "0.001_1e-05" = expression(atop("E(s) = 0.001,", paste(delta[C], "  = 1e-5"))), 
                            "0.1_1e-07"  = expression(atop("E(s) = 0.1,", paste(delta[C], "  = 1e-7"))),
                            "0.1_1e-05"  = expression(atop("E(s) = 0.1,", paste(delta[C], "  = 1e-5"))))) 


b = multi_data %>%
  select(obs_set, shape_real, scale_real, es, mu_real, 
         mean_corr_ppc, corr95_low_ppc, corr95_hi_ppc) %>%
  mutate(type = "multi") %>%
  bind_rows(single_data %>%
              rename(shape_real = alpha_real, scale_real = beta_real) %>%
              mutate(es = shape_real*scale_real, type = "single") %>%
              select(obs_set, shape_real, scale_real, es, mu_real,
                     mean_corr_ppc, corr95_low_ppc, corr95_hi_ppc, type)) %>%
  mutate(sim_params = paste(es, mu_real, sep = "_")) %>%
  ggplot(aes(sim_params, mean_corr_ppc, color = type)) +
  geom_pointrange(aes(ymin = corr95_low_ppc,ymax = corr95_hi_ppc), alpha=0.5,
                  position=position_dodge2(width=0.75)) +
  facet_wrap(~shape_real) +
  ylab("Posterior predictive check \ncorrelation") +
  scale_color_manual(labels = c("Directly inferred MAP distribution",
                                  "Distribution from single observation MAPs"), name = "",
                     #values=c("#453781", "#1f968b", "#75d054")) +
                     values=c("blue", "darkgreen")) +
  ggtitle("B") +
  scale_x_discrete(name = "",
                   labels=c("0.001_1e-07"  = expression(atop("E(s) = 0.001,", paste(delta[C], "  = 1e-7"))),
                            "0.001_1e-05" = expression(atop("E(s) = 0.001,", paste(delta[C], "  = 1e-5"))), 
                            "0.1_1e-07"  = expression(atop("E(s) = 0.1,", paste(delta[C], "  = 1e-7"))),
                            "0.1_1e-05"  = expression(atop("E(s) = 0.1,", paste(delta[C], "  = 1e-5"))))) 

sup = a/b
sup
ggsave(paste0(fig_path,"SupplementaryFig10.pdf"), plot = sup, width = 20, height = 10, units = "in")
ggsave(paste0(fig_path,"SupplementaryFig10.png"), plot = sup, width = 20, height = 10, units = "in")
```
We assessed the performance of inference from a set of observations using out-of-sample posterior predictive accuracy [94]that inferring  and  from a set of observations results in lower posterior predictive accuracy compared to inferring sC from a single observation (Supplementary File x, Supplementary Figure 7)

