---
title: "Figure 4 and associated supplementary figures"
output: html_notebook
author: Grace Avecilla
---

What is the best model? WF is good enough - and faster and generalizable (NPE)  

```{r libs and functions}
library(tidyverse)
library(patchwork)
library(scales)
library(ggbeeswarm)
theme_set(theme_classic(base_size = 18))

read_est_params = function(x) {
  if(str_detect(x, "NPE")==TRUE){
    file = read_csv(x, col_names = c("Model","Method","col_to_sep",
                                     "s_est","mu_real",
                                     "mu_est","s_snv","m_snv","rmse_map",
                                     "mean_rmse_ppc","rmse95_low_ppc","rmse95_hi_ppc",
                                     "corr_map",
                                     "mean_corr_ppc","corr95_low_ppc", "corr95_hi_ppc",
                                     "fit_95hdi_low","fit_95hdi_high",
                                     "mut_95hdi_low","mut_95hdi_high",
                                     "aic", "dic", "waic1", "waic2")) %>%
      separate(col_to_sep, into = c(NA,NA,NA,"n_presim", "presim_rep","s_real"), 
               sep="([_.])") %>%
      mutate(s_real = str_remove(s_real, "csv")) %>%
      mutate(s_est_10 = s_est, s_real_10 = as.numeric(s_real),
             mu_est_10 = mu_est, mu_real_10 = mu_real,
             s_est = 10^s_est, s_real = 10^as.numeric(s_real),
             mu_est = 10^mu_est, mu_real = 10^mu_real) %>%
      mutate(mu_ratio = mu_est/mu_real, s_ratio = s_est/s_real,
             mu_diff = mu_est - mu_real, s_diff = s_est - s_real) %>%
      mutate(flow_type = if_else(str_detect(x, "nsf"), "NSF", "MAF"))
  }
  if(str_detect(x, "ABC")==TRUE) {
    file = read_csv(x, col_names = c("Model","Method","col_to_sep",
                                     "s_est","mu_real",
                                     "mu_est","s_snv","m_snv","rmse_map",
                                     "mean_rmse_ppc","rmse95_low_ppc","rmse95_hi_ppc",
                                     "corr_map",
                                     "mean_corr_ppc","corr95_low_ppc", "corr95_hi_ppc",
                                     "fit_95hdi_low","fit_95hdi_high",
                                     "mut_95hdi_low","mut_95hdi_high",
                                     "aic", "dic", "waic1", "waic2")) %>%
      separate(col_to_sep, into = c("n_presim","s_real"), 
               sep="([-])") %>%
      mutate(s_real = as.numeric(s_real)*-1) %>%
      mutate(presim_rep = case_when(str_detect(x, "_1_") ~ "1",
                                    str_detect(x, "_2_") ~ "2",
                                    str_detect(x, "_3_") ~ "3")) %>%
      select("Model","Method","n_presim", "presim_rep","s_real",
             "s_est","mu_real",
             "mu_est","s_snv","m_snv","rmse_map",
             "mean_rmse_ppc","rmse95_low_ppc","rmse95_hi_ppc",
             "corr_map",
             "mean_corr_ppc","corr95_low_ppc", "corr95_hi_ppc",
             "fit_95hdi_low","fit_95hdi_high",
             "mut_95hdi_low","mut_95hdi_high",
             "aic", "dic", "waic1", "waic2") %>%
      mutate(s_est_10 = s_est, s_real_10 = as.numeric(s_real),
             mu_est_10 = mu_est, mu_real_10 = mu_real,
             s_est = 10^s_est, s_real = 10^as.numeric(s_real),
             mu_est = 10^mu_est, mu_real = 10^mu_real) %>%
      mutate(mu_ratio = mu_est/mu_real, s_ratio = s_est/s_real,
             mu_diff = mu_est - mu_real, s_diff = s_est - s_real) %>%
      mutate(flow_type = NA) %>%
      mutate(n_presim = if_else(n_presim == "100", "10000", "100000"))
  }
  return(file)
}

read_cross_sim = function(x) {
  data = read_est_params(x)
  data = data %>% 
    mutate(obs_model = case_when(str_detect(x,"Chemo_obs") ~ "Chemo",
                                 str_detect(x, "WF_obs") ~ "WF",
                                 !str_detect(x,"obs") ~ Model))
  return(data)
}
```


```{r get data}
file_dir = "/Volumes/GoogleDrive/.shortcut-targets-by-id/19NQRq-XLPpjqJGmLbP4MHRjala2WVahk/Avecilla et al Simulating CNV evolution/scripts/cnv_sims_inference/hpc_output/test_performance/single/param_estimates/"
files1 = list.files(path = file_dir,
                    pattern='*est_real_params.csv')
files1 = paste0(file_dir,files1)

file_dir = "/Volumes/GoogleDrive/.shortcut-targets-by-id/19NQRq-XLPpjqJGmLbP4MHRjala2WVahk/Avecilla et al Simulating CNV evolution/scripts/cnv_sims_inference/hpc_output/test_performance/cross_simulation/"
files = list.files(path = paste0(file_dir, "param_estimates/"), pattern = '*est_real_params.csv')
files = paste0(file_dir,"param_estimates/", files)

files = c(files,files1)

data_list = map(files,read_cross_sim)
data_all = do.call(rbind, data_list) %>%
  mutate(Method = str_replace(Method, "SNPE", "NPE")) %>%
  mutate(relative_error_mut = (mu_real - mu_est)/mu_real, relative_error_s = (s_real - s_est)/s_real)

# percent in HDR, counted manually
hdr_data = read_csv(paste0(file_dir, "HDR_cross_sim.csv"))


fig_path = "/Volumes/GoogleDrive/.shortcut-targets-by-id/19NQRq-XLPpjqJGmLbP4MHRjala2WVahk/Avecilla et al Simulating CNV evolution/figures/"
```

# Fig 4

```{r set up for Fig4}
data = data_all %>% filter(n_presim == "100000" & flow_type == "NSF") %>% 
  unite(mod_method, Model, Method, sep = " ", remove=F) %>%
  unite(sim_params,s_real, mu_real, remove = F) %>%
  mutate(obs_model = if_else(obs_model == "Chemo", "Chemo observation", "WF observation"))

data$sim_params = factor(data$sim_params, levels = c("0.001_1e-05", "0.001_1e-07", "0.1_1e-05", "0.1_1e-07"),
                         ordered = TRUE, labels=c(expression(atop(paste(italic("s"[C])," = 0.001,"), paste(italic(delta[C]), "  = 1e-5"))),
                                                  expression(atop(paste(italic("s"[C])," = 0.001,"), paste(italic(delta[C]), "  = 1e-7"))),
                                                  "0.1_1e-05"  = expression(atop(paste(italic("s"[C])," = 0.1,"), paste(italic(delta[C]), "  = 1e-5"))),
                                                  "0.1_1e-07"  = expression(atop(paste(italic("s"[C])," = 0.1,"), paste(italic(delta[C]), "  = 1e-7")))))


color_vec = c("#4daf4a", "#984ea3","#e41a1c", "#377eb8", "black")
mod_meth = c("Chemo NPE", "WF NPE", "Chemo ABC-SMC", "WF ABC-SMC", "Pseudo-observation")

#labs <- c("s relative error", expression(paste(delta," relative error")), "Total relative error")
#names(labs) <- c("relative_error_s", "relative_error_mut", "total_rel_error")

hdr_data_long = hdr_data %>%
  filter(n_presim == 100000 & flow_type == "NSF") %>%
  unite(mod_method, infer_model, method, sep = " ", remove=F) %>%
  pivot_longer(cols = starts_with("0."), names_to = "sim_params", values_to = "in_hdr") %>% 
  mutate(in_hdr_prop = in_hdr/5) %>%
  mutate(flow_type = if_else(is.na(flow_type), "", flow_type)) %>%
  mutate(obs_model = if_else(obs_model == "Chemo", "Chemo observation", "WF observation"))

```



```{r a}
data4 = data %>%
  pivot_longer(cols = c("mu_ratio", "s_ratio"), 
               names_to = "error_type", values_to = "log10 ( MAP parameter / true parameter )") %>%
  mutate(`log10 ( MAP parameter / true parameter )` = log10(`log10 ( MAP parameter / true parameter )`)) %>%
  filter(obs_model == "Chemo observation") %>%
  mutate(obs_model = str_replace(obs_model, " ","~")) 

a = data4 %>%
  filter(s_real == 0.001 & mu_real == 1e-5) %>%
  ggplot(aes(error_type,`log10 ( MAP parameter / true parameter )`, color=mod_method)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge2(width=1))+
  scale_color_manual(values=color_vec, 
                     name="Inference model\nand method",
                     breaks=mod_meth) +
  geom_beeswarm(aes(fill=mod_method, shape = as.factor(presim_rep)), dodge.width = .75, alpha = 0.5) +
  scale_shape_manual(values = c(21, 22, 23), name = "Training set") +
  scale_fill_manual(values=color_vec, 
                    name="",
                    limits=mod_meth, guide = 'none') +
  ylab("log10 ( MAP / true )") +
  xlab("") +
  facet_wrap(~sim_params, labeller=label_parsed, scales = "free") +
  scale_x_discrete(labels=c("s_ratio"  = expression(italic("s"[C])),
                            "mu_ratio" = expression(italic(delta[C])))) + 
                            #"total_rel_error"  = "total")) +
  #xlab("Relative error type") +
  theme(panel.spacing.y=unit(1, "lines"),panel.spacing.x=unit(1, "lines")) +
  ggtitle("A") +
  geom_hline(yintercept = 0, alpha = 0.5) +
  theme(legend.position = "none") +
  geom_hline(yintercept = c(-1,1), alpha = 1, linetype = "dotted")
a
```

```{r b}
b = data4 %>%
  filter(s_real == 0.001 & mu_real == 1e-7) %>%
  ggplot(aes(error_type,`log10 ( MAP parameter / true parameter )`, color=mod_method)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge2(width=1))+
  scale_color_manual(values=color_vec, 
                     name="Inference model\nand method",
                     breaks=mod_meth) +
  geom_beeswarm(aes(fill=mod_method, shape = as.factor(presim_rep)), dodge.width = .75, alpha = 0.5) +
  scale_shape_manual(values = c(21, 22, 23), name = "Training set") +
  scale_fill_manual(values=color_vec, 
                    name="",
                    limits=mod_meth, guide = 'none') +
  xlab("") +
  ylab("") +
  facet_wrap(~sim_params, labeller=label_parsed, scales = "free") +
  scale_x_discrete(labels=c("s_ratio"  = expression(italic("s"[C])),
                            "mu_ratio" = expression(italic(delta[C])))) + 
                            #"total_rel_error"  = "total")) +
  #xlab("Relative error type") +
  theme(panel.spacing.y=unit(1, "lines"),panel.spacing.x=unit(1, "lines")) +
  ggtitle("B") +
  geom_hline(yintercept = 0, alpha = 0.5) +
  theme(legend.position = "none") +
  geom_hline(yintercept = c(-1,1), alpha = 1, linetype = "dotted")
b
```
```{r c}
c = data4 %>%
  filter(s_real == 0.1 & mu_real == 1e-5) %>%
  ggplot(aes(error_type,`log10 ( MAP parameter / true parameter )`, color=mod_method)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge2(width=1))+
  scale_color_manual(values=color_vec, 
                     name="Inference model\nand method",
                     breaks=mod_meth) +
  geom_beeswarm(aes(fill=mod_method, shape = as.factor(presim_rep)), dodge.width = .75, alpha = 0.5) +
  scale_shape_manual(values = c(21, 22, 23), name = "Training set") +
  scale_fill_manual(values=color_vec, 
                    name="",
                    limits=mod_meth, guide = 'none') +
  xlab("") +
  ylab("") +
  facet_wrap(~sim_params, labeller=label_parsed, scales = "free") +
  scale_x_discrete(labels=c("s_ratio"  = expression(italic("s"[C])),
                            "mu_ratio" = expression(italic(delta[C])))) + 
                            #"total_rel_error"  = "total")) +
  #xlab("Relative error type") +
  theme(panel.spacing.y=unit(1, "lines"),panel.spacing.x=unit(1, "lines")) +
  ggtitle("C") +
  geom_hline(yintercept = 0, alpha = 0.5) +
  theme(legend.position = "none") +
  geom_hline(yintercept = c(-1,1), alpha = 1, linetype = "dotted")
c
```

```{r d}
d = data4 %>%
  filter(s_real == 0.1 & mu_real == 1e-7) %>%
  ggplot(aes(error_type,`log10 ( MAP parameter / true parameter )`, color=mod_method)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge2(width=1))+
  scale_color_manual(values=color_vec, 
                     name="Model\nand method",
                     breaks=mod_meth) +
  geom_beeswarm(aes(fill=mod_method, shape = as.factor(presim_rep)), dodge.width = .75, alpha = 0.5) +
  scale_shape_manual(values = c(21, 22, 23), name = "Training set") +
  scale_fill_manual(values=color_vec, 
                    name="",
                    limits=mod_meth, guide = 'none') +
  xlab("") +
  ylab("") +
  facet_wrap(~sim_params, labeller=label_parsed, scales = "free") +
  scale_x_discrete(labels=c("s_ratio"  = expression(italic("s"[C])),
                            "mu_ratio" = expression(italic(delta[C])))) + 
                            #"total_rel_error"  = "total")) +
  #xlab("Relative error type") +
  theme(panel.spacing.y=unit(1, "lines"),panel.spacing.x=unit(1, "lines")) +
  ggtitle("D") +
  geom_hline(yintercept = 0, alpha = 0.5) +
  geom_hline(yintercept = c(-1,1), alpha = 1, linetype = "dotted")
d
```


```{r Figure 4}
layout <- "
ABCD
"
fig4 = a + b + c + d + 
  plot_layout(design = layout)
fig4
ggsave(paste0(fig_path,"Figure4.pdf"), plot = fig4, width = 18, height = 6, units = "in")
ggsave(paste0(fig_path,"Figure4.png"), plot = fig4, width = 18, height = 6, units = "in")

```
*Figure 4* These show the results of inference on five simulated synthetic observations generated using either the Wright-Fisher (WF) or chemostat (Chemo) model per combination of fitness effect s and mutation rate 𝛿C. We performed inference on each synthetic observation using either the Wright-Fisher (WF) or chemostat (Chemo) model. For NPE, each training set corresponds to an independent amortized posterior trained with 100,000 simulations, with which each synthetic observation was evaluated. Boxplots and points show the ratio of MAP estimate to true parameters as compared to true parameters, log transformed, for s and  δC.

```{r sup fig 8}
s_a = hdr_data_long %>%
  group_by(obs_model,mod_method, sim_params, HDR) %>%
  summarize(mean_in_hdr_prop = mean(in_hdr_prop)) %>%#*100) %>%
  filter(HDR == "50%") %>%
  ggplot(aes(sim_params, mean_in_hdr_prop,fill=mod_method)) +
  geom_bar(stat="identity", position="dodge") +
  geom_point(data=hdr_data_long %>%
               filter(method == "NPE" & flow_type != "MAF") %>%
               filter(n_presim == 100000) %>%
               filter(HDR=="50%") %>%
               mutate(in_hdr_prop = in_hdr/5), 
             aes(sim_params, in_hdr_prop,fill=mod_method,shape = as.factor(rep)),
             color="black",
             position = position_dodge2(width = 1),
             size=3) +
  scale_fill_manual(values=color_vec,name="Inference model\nand method", breaks=mod_meth) +
  scale_shape_manual(values = c(21, 22, 23), name = "Training set") +
  xlab("") +
  ylab("Percent of true parameters\nin 50% HDR") +
  scale_y_continuous(labels = scales::percent) +#, limits = c(0,100)) +
  ggtitle("A") +
  scale_x_discrete(labels=c("0.001_1e-7"  = expression(atop(paste(italic("s"[C])," = 0.001,"), paste(italic(delta[C]), "  = 1e-7"))),
                                                                                  "0.001_1e-5" = expression(atop(paste(italic("s"[C])," = 0.001,"), paste(italic(delta[C]), "  = 1e-5"))), 
                                                                                  "0.1_1e-7"  = expression(atop(paste(italic("s"[C])," = 0.1,"), paste(italic(delta[C]), "  = 1e-7"))),
                                                                                  "0.1_1e-5"  = expression(atop(paste(italic("s"[C])," = 0.1,"), paste(italic(delta[C]), "  = 1e-5"))))) +
  facet_wrap(~obs_model)



s_b = hdr_data_long %>%
  filter(method == "NPE" & flow_type != "MAF") %>%
  filter(n_presim == 100000) %>%
  group_by(obs_model,mod_method, sim_params, HDR) %>%
  summarize(mean_in_hdr_prop = mean(in_hdr_prop)) %>%#*100) %>%
  filter(HDR == "95%") %>%
  ggplot(aes(sim_params, mean_in_hdr_prop,fill=mod_method)) +#, shape = as.factor(rep))) +
  #geom_boxplot() +
  geom_bar(stat="identity", position="dodge") +
  geom_point(data=hdr_data_long %>%
               filter(method == "NPE" & flow_type != "MAF") %>%
               filter(n_presim == 100000) %>%
               filter(HDR=="95%") %>%
               mutate(in_hdr_prop = in_hdr/5), 
             aes(sim_params, in_hdr_prop,fill=mod_method,shape = as.factor(rep)),
             color="black",
             position = position_dodge2(width = 1),
             size=3) +
  scale_fill_manual(values=color_vec,name="Inference model\nand method", breaks=mod_meth) +
  scale_shape_manual(values = c(21, 22, 23), name = "Training set") +
  xlab("") +
  ylab("Percent of true parameters\nin 95% HDR") +
  scale_y_continuous(labels = scales::percent) +#, limits = c(0,100)) +
  ggtitle("B") +
  scale_x_discrete(labels=c("0.001_1e-7"  = expression(atop(paste(italic("s"[C])," = 0.001,"), paste(italic(delta[C]), "  = 1e-7"))),
                                                                                  "0.001_1e-5" = expression(atop(paste(italic("s"[C])," = 0.001,"), paste(italic(delta[C]), "  = 1e-5"))), 
                                                                                  "0.1_1e-7"  = expression(atop(paste(italic("s"[C])," = 0.1,"), paste(italic(delta[C]), "  = 1e-7"))),
                                                                                  "0.1_1e-5"  = expression(atop(paste(italic("s"[C])," = 0.1,"), paste(italic(delta[C]), "  = 1e-5"))))) +
  facet_wrap(~obs_model)

supp = s_a / s_b
supp
ggsave(paste0(fig_path,"SupplementaryFig8.pdf"), plot = supp, width = 18, height = 15, units = "in")
ggsave(paste0(fig_path,"SupplementaryFig8.png"), plot = supp, width = 18, height = 15, units = "in")

```
*Supplemental Figure 8* These show the results of inference on five simulated synthetic observations generated using either the Wright-Fisher (WF) or chemostat (Chemo) model per combination of fitness effect s and mutation rate 𝛿C. We performed inference on each synthetic observation using both models. For NPE, each training set corresponds to an independent amortized posterior trained with 100,000 simulations, with which each synthetic observation was evaluated. A) Percentage of true parameters within the 50% HDR. The bar height shows the average of three training sets. B) Percentage of true parameters within the 95% HDR. The bar height shows the average of three training sets.

